---
title: "Spatial Regression of Stormwater Pollutants"
author: "Christian Nilsen"
output:
  html_document:
    s8data_print: paged
---
#To do: 
*add bibilography
*select a mixed effects model

#Introduction
This notebook demonstrates spatial regression of stormwater pollutant data for Puget Sound. It combines stormwater outfall monitoring data with spatial regression to predict pollutant concentrations across Puget Sound. 

The stormwater outfall data is available from the Department of Ecology at
https://data.wa.gov/Natural-Resources-Environment/Municipal-Stormwater-Permit-Outfall-Data/d958-q2ci
It uses the Socrata REST api in the form of 

https://data.wa.gov/resource/d958-q2ci.json?parameter=Copper%20-%20Water%20-%20Total




```{r echo=FALSE}
library(RSocrata)
if (!require('dplyr')) install.packages('dplyr')
if (!require('NADA')) install.packages('NADA')
if (!require('RSocrata')) install.packages('RSocrata')
if (!require('Hmisc')) install.packages('Hmisc')
if (!require('tidyverse')) install.packages('tidyverse')
if (!require('formattable')) install.packages('formattable')
if (!require('kableExtra')) install.packages('kableExtra')
if (!require('survival')) install.packages('survival')

if (!require('PerformanceAnalytics')) install.packages("PerformanceAnalytics")
library(NADA)
library(RSocrata)
library(dplyr)
library(Hmisc)
library(formattable)
library(survminer)
library(tidyverse)
library(kableExtra)
library(survival)
library(PerformanceAnalytics)
```
Get data
```{r}
url <- ("https://data.wa.gov/resource/rc6b-fvgb.json")
parameter <- 'Copper - Water - Total'
#construct the SQL call 
#apiCall <- paste0(url,"?$where=Result_Data_Qualifier<>'REJ' AND parameter= ","'",parameter,"'")
apiCall <- paste0(url,"?$where=parameter= ","'",parameter,"'")
data <-as_tibble(read.socrata(apiCall))
data <- (filter(data,!result_data_qualifier %in% 'REJ'))
data <- type.convert(data)

```
Looks OK

```{r echo=TRUE}
#read in spatial data and join
s8data <- data %>% 
  dplyr::select(location_id,study_name, parameter,season,new_result_value,nondetect_flag, new_result_units,field_collection_start_date) #sampling data
luLookup <-
 (read.csv("Data/luLookup.csv",  na = "NA")) #land use data
eeIndicies <-
  read.csv("Data/eeIndiciesNorm.csv") #spatial parameters
s8xy <-
  read.csv("Data/s8xy.csv") #xy coordinates of sampling locations

#execute join and do some fomatting
names(s8data)[1] <- "Location"
s8data <- s8data %>%
    merge(eeIndicies, by = "Location") %>%
    merge(luLookup, by = "Location") %>%
    merge(s8xy, by = "Location") 
#s8data <- as_tibble(s8data)
s8data$Year <- with(s8data, format(as.Date(field_collection_start_date),"%Y"))

#Trim location names so that Port of Seattle outfall locations are the same (all same study)
s8data$Location <- substr(s8data$Location, 1, 12)
#Plot data to view
plot.title <- paste0(s8data$parameter, ' (',s8data$new_result_units,')')
plot <- ggplot(data = s8data) + 
  geom_point(mapping = aes(x = 1, y = new_result_value, color = study_name), position = 'jitter')  
plot+ggtitle(plot.title)
```
Looks good. Lets look at how concentration are correleated with landuse. 



```{r}
s8data$Loc.Year <- paste0(s8data$Location, '-', s8data$Year)
# Box plot with dot plot
plot <- ggplot(s8data, aes(Year, new_result_value, fill = Loc.Year)) + 
  geom_boxplot()+theme(legend.position = "none")+scale_y_log10()+ggtitle(plot.title)
 
   
plot
```
Since the data set has mixed detection limits, we will use the kaplan-meier estimator to evaluate median values
```{r echo=FALSE}
#first determine which are censored and which are not 
obs      <- as.numeric(s8data$new_result_value)
censored <-as.logical(s8data$nondetect_flag)
groups = as.factor(s8data$Location)
KM = cenfit(obs, censored, groups)
b <- as.data.frame(median(KM))
KM.medians <- tibble(site = row.names(b), median = (b$`median(KM)`))
kable(KM.medians)

```


Look at landuses 
```{r echo=TRUE}
cor.data <- (cbind(s8data$new_result_value, s8data[, 17:21])) #land uses
chart.Correlation(cor.data, histogram=FALSE, method = 'pearson')
cor.matrix <- rcorr(as.matrix(cor.data),s8data$nondetect_flag)
r<- (cor.matrix$r)[,1]
P<- (cor.matrix$P)[,1]
table <- (cbind(r,P))
kable(table)

```
Look at other indices 
```{r}
cor.data <- cbind(s8data$new_result_value, s8data[, 10:16], s8data$age, s8data$elev) #indicies from earth engine
chart.Correlation(cor.data, histogram=FALSE, method = 'pearson')
cor.matrix <- rcorr(as.matrix(cor.data),s8data$nondetect_flag)
```

Imperviousness is the highest correleated, followed by elevation, then age and ndvi mean. 

Let's choose 0.25 as the cutoff for correlation. This gives the following as variables to use: 
*Landuse 
  *Percent COM
  *Percent SFR
*Indicies
  *imperviousness
  *ndbi
  *roads
```{r}
r<- (cor.matrix$r)[,1]
P<- (cor.matrix$P)[,1]
table <- (cbind(r,p))
kable(table)
```

Begin linear model  (ignore non-detects for now)
```{r}
library(nlme)
fit1 <- gls(new_result_value ~ COM + SFR + Impee + ndbi +roads, data=s8data,method="ML")
summary(fit1)


```
Begin mixed effect model
```{r}
#try loc and year as grouping factors
fit2 <- lme(new_result_value ~ COM + SFR + Impee + ndbi +roads, data=s8data, random = ~1|Loc.Year,method="ML")  
summary(fit2)
```
Display ANOVA table
```{r}
fit3 <- lme(new_result_value ~ Impee  +roads, data=s8data, random = ~1|Loc.Year,method="ML") 
summary(fit3)
anova(fit1,fit2)
```

Fit2 has a slightly lower AIC Begin forward selection 

```{r}
library(leaps)
library(AICcmodavg)

Cand.models<-list()
Cand.models[[1]] <- fit2
Cand.models[[2]] <- fit3


#create a vector of names to trace back models in set
Modnames<-paste("mod", 1:length(Cand.models), sep="")

#generate AICc table
bictab(cand.set = Cand.models, modnames = Modnames, sort=TRUE)
plot(fit1)

```
