---
title: "TotalPAH"
output: html_document
---

---
#To do: 
*add in rainfall 
*clean up eeIncidies
*add more exploratory section
*follow Tufts example
*add bibilography
*select a mixed effects model

#Introduction
This notebook demonstrates spatial regression of stormwater pollutant data for Puget Sound. It combines stormwater outfall monitoring data with spatial regression to predict pollutant concentrations across Puget Sound. 

The stormwater outfall data is available from the Department of Ecology at
https://data.wa.gov/Natural-Resources-Environment/Municipal-Stormwater-Permit-Outfall-Data/d958-q2ci
It uses the Socrata REST api in the form of 

https://data.wa.gov/resource/d958-q2ci.json?parameter=Copper%20-%20Water%20-%20Total




```{r include=FALSE}
library(RSocrata)

if (!require('NADA')) install.packages('NADA')
if (!require('RSocrata')) install.packages('RSocrata')
if (!require('Hmisc')) install.packages('Hmisc')
if (!require('tidyverse')) install.packages('tidyverse')
if (!require('formattable')) install.packages('formattable')
if (!require('kableExtra')) install.packages('kableExtra')
if (!require('survival')) install.packages('survival')

if (!require('PerformanceAnalytics')) install.packages("PerformanceAnalytics")
library(NADA)
library(RSocrata)
library(dplyr)
library(Hmisc)
library(formattable)
library(survminer)
library(tidyverse)
library(kableExtra)
library(survival)
library(PerformanceAnalytics)
```
# Get data and plot it
```{r}
url <- ("https://data.wa.gov/resource/rc6b-fvgb.json")
parameter <- 'Total PAH - Water - Total'
#construct the SQL call 
#apiCall <- paste0(url,"?$where=Result_Data_Qualifier<>'REJ' AND parameter= ","'",parameter,"'")
apiCall <- paste0(url,"?$where=parameter= ","'",parameter,"'")
data <-as_tibble(read.socrata(apiCall))
data <- (filter(data,!result_data_qualifier %in% 'REJ'))
data <- type.convert(data)

```
Looks OK

```{r echo=TRUE}
#read in spatial data and join


s8data <- data %>% 
   dplyr::select(location_id,study_id, parameter,season,new_result_value,nondetect_flag, new_result_units,field_collection_start_date,type) #sampling data


luLookup <-
 (read.csv("Data/luLookup.csv",  na = "NA")) #land use data
eeIndicies <-
  read.csv("Data/eeIndiciesNorm.csv") #spatial parameters
s8xy <-
  read.csv("Data/s8xy.csv") #xy coordinates of sampling locations


colnames(s8data)[colnames(s8data)=="location_id"] <- "Location"
```


```{r echo=TRUE}
s8data <- s8data %>%
    merge(eeIndicies, by = "Location") %>%
    merge(luLookup, by = "Location") %>%
    merge(s8xy, by = "Location") 
#s8data <- as_tibble(s8data)
s8data$Year <- with(s8data, format(as.Date(field_collection_start_date),"%Y"))

#Trim location names so that Port of Seattle outfall locations are the same (all same study)
s8data$Location <- substr(s8data$Location, 1, 12)
#Plot data to view

# Plot by study 
plot.title <- paste0(s8data$parameter, ' (',s8data$new_result_units,')')

plot <- ggplot(s8data, aes(study_id,new_result_value, fill = study_id)) + 
  geom_boxplot()+theme(legend.position = "none")+scale_y_log10()+ggtitle(plot.title)
 
   
plot
```
#Explore the data more 
```{r}

# Plot by type 
plot <- ggplot(s8data, aes(type,new_result_value, fill = study_id)) + 
  geom_boxplot()+theme(legend.position = "none")+scale_y_log10()+ggtitle(plot.title)
 
   
plot
```
#Explore Correleation
Look at landuses 
```{r echo=TRUE}
cor.data <- (cbind(s8data$new_result_value, s8data[, 18:22])) #land uses
chart.Correlation(cor.data, histogram=FALSE, method = 'pearson')
cor.matrix <- rcorr(as.matrix(cor.data),s8data$nondetect_flag)
r<- (cor.matrix$r)[,1]
P<- (cor.matrix$P)[,1]
table <- (cbind(r,P))
kable(table)

```
Look at other indices 
```{r}
cor.data <- cbind(s8data$new_result_value, s8data[, 10:17], s8data$age, s8data$elev) #indicies from earth engine
chart.Correlation(cor.data, histogram=FALSE, method = 'pearson')
cor.matrix <- rcorr(as.matrix(cor.data),s8data$nondetect_flag)
```

It looks like outliers are messing things up. 

```{r}
#plot everything as one boxplot to view outliers 

boxplot(s8data$new_result_value)$out


```

```{r}
#normally it is a good idea to keep outliers, but that one is so much larger, this is probably skewing things. 

s8data<-
  s8data [order(-s8data$new_result_value),] 
s8data <-  s8data[-1,]

#replot 
boxplot(s8data$new_result_value)$out
```
Better! Try correlation again 
This time log transform the data 

```{r echo=TRUE}
cor.data <- (cbind(log(s8data$new_result_value), s8data[, 18:22])) #land uses
chart.Correlation(cor.data, histogram=FALSE, method = 'pearson')
cor.matrix <- rcorr(as.matrix(cor.data),s8data$nondetect_flag)
r<- (cor.matrix$r)[,1]
P<- (cor.matrix$P)[,1]
table <- (cbind(r,P))
kable(table)
```

```{r}

cor.data <- cbind(log(s8data$new_result_value), s8data[, 10:17], s8data$age, s8data$elev) #indicies from earth engine
chart.Correlation(cor.data, histogram=FALSE, method = 'pearson')
cor.matrix <- rcorr(as.matrix(cor.data),s8data$nondetect_flag)


```


```{r}
r<- (cor.matrix$r)[,1]
P<- (cor.matrix$P)[,1]
table <- (cbind(r,P))
(table)
```
```{r}


# All Subsets Regression
library(leaps)


leaps<-regsubsets(new_result_value ~ 
                    MFR + IND + ghm + pm25 + ndbi +Impee + elev + roofs + elev, data=s8data,method = "forward")
# view results 

# plot a table of models showing variables in each model.
# models are ordered by the selection statistic.
#layout(matrix(c(1, 2), 2, 1)) # optional 2 graphs/page
plot(leaps,scale="adjr2")
plot(leaps,scale="bic")
```
We will pass IND, Impee, roofs,  to our mixed model

#Let's see if there is a distribution we can use

```{r}
library(MASS)
library(qualityTools)
#normal distribution
x <- s8data$new_result_value
qqPlot(x, "log-normal")
qqPlot(x, "weibull")
#qqPlot(x, "gamma")
qqPlot(x, "normal")
qqPlot(x, 'exponential')
qqPlot(log(x+1), "weibull")#this is the Gumbell distribution


```
Data are not normally distributed, mixed effects modeling cannot have normal distribution assumptions for random effects. 

assume a weibell or gamma distribution for mixed effects. These are both forms of the exponetional function

#Generalized Mixed Model 
PQL <- glmmPQL(Aggression.t ~ Relation + Season, ~1 | Observer/Test.ID, family = gaussian(link = "log"),
    data = recog, verbose = FALSE)
    
```{r}
PQL <- glmmPQL(new_result_value ~ Impee+roofs+IND, data=s8data, random = ~1|Location,family = gaussian(link='log'))



summary(PQL)
```
None are significant



    
#Dealing with Censored Data
Since the data set has mixed detection limits, we will use the kaplan-meier estimator to evaluate median values